---
title: "Electoral Commission donations data"
format: gfm
---

The Electoral Commission [publishes data](https://search.electoralcommission.org.uk/?currentPage=1&rows=100&sort=AcceptedDate&order=desc&tab=1&et=pp&et=ppm&et=tp&et=perpar&et=rd&isIrishSourceYes=true&isIrishSourceNo=true&prePoll=false&postPoll=true&register=gb&register=ni&register=none&optCols=Register&optCols=CampaigningName&optCols=AccountingUnitsAsCentralParty&optCols=IsSponsorship&optCols=IsIrishSource&optCols=RegulatedDoneeType&optCols=CompanyRegistrationNumber&optCols=Postcode&optCols=NatureOfDonation&optCols=PurposeOfVisit&optCols=DonationAction&optCols=ReportedDate&optCols=IsReportedPrePoll&optCols=ReportingPeriodName&optCols=IsBequest&optCols=IsAggregation) on donations to political parties in the UK. This data contains some information about the donor, such as their name and status, but not much else.

This example shows how to use Augmenta to enrich this data with additional information about the donors, such as their industry, location, and size. This can help to identify patterns in donations, such as which industries are more likely to donate to which parties.

## Data prep

```{r}
#| output: false
library(tidyverse)
```

First, let's download and read the data.

```{r}
#| output: false
donations_raw <- read_csv("https://search.electoralcommission.org.uk/api/csv/Donations?start={start}&rows=100&query=&sort=AcceptedDate&order=desc&et=pp&date=Accepted&from=2024-07-01&to=2024-09-30&rptPd=&prePoll=false&postPoll=true&register=ni&register=gb&isIrishSourceYes=true&isIrishSourceNo=true&includeOutsideSection75=true")
```

For the purposes of this example, we'll only keep the columns that contain information about the donor. We'll also only look at the first 10 rows to keep things simple.

```{r}
#| output: false
donations <- donations_raw |>
  slice_head(n = 10) |>
  select(RegulatedEntityName, DonorName, DonorStatus, CompanyRegistrationNumber, Value) |>
  mutate(Value = parse_number(Value)) |>
  # remove duplicate rows
  group_by(RegulatedEntityName, DonorName, DonorStatus, CompanyRegistrationNumber) |>
  summarise(Value = sum(Value)) |>
  ungroup()
```

Finally, let's save this data as a CSV file.

```{r}
write_csv(donations, "data/donations.csv")
```

## Augmentation

Now that we have our data, we can use Augmenta to enrich it with additional information about the donors.

## Setup

First, we need to configure our project. Create a new directory for you project. In it, create a file called `config.yaml` and paste this into it:

```yaml
input_csv: data/donations.csv
output_csv: data/donations_classified.csv
model:
  name: openai/gpt-4o-mini
  max_tokens: 20000
query_col: DonorName
search:
  engine: brave
  country: GB
  results: 10
  rate_limit: 2
prompt:
  system: You are an expert researcher whose job is to classify individuals and companies based on their industry.
  user: |
    # Instructions

    The following documents are web search results about {{DonorName}}, a {{RegulatedEntityName}} donor identified in data published by the UK Electoral Commission. Your task is to determine what industry {{DonorName}} belongs to. The documents could be about a company, a trade group, a union, an individual, etc.
    
    If {{DonorName}} is an individual, you should classify them based on their profession or the industry they are closest associated with. If the documents are about multiple individuals, or if it's not clear which individual the documents refer to, please set the industry to "Don't know" and the confidence level to 1. For example, there's no way to know for certain that someone named "John Smith" in the documents is the same person as the donor in the Electoral Commission.

    We also know that the donor is a {{DonorStatus}}.

    Use the information provided in the documents to make your decision. Be critical, use common sense and respond only in English. Now, please proceed with your analysis and classification of {{DonorName}}.    
structure:
  chain_of_thought:
    type: str
    description: Understand the task and make a complete plan to accomplish it. Explain your reasoning. Then, carry out the plan, think of potential issues that might come up, then show your answer. Assess your own answer, think of ways to improve it, and then show the improved answer.
  industry:
    type: str
    description: What industry is this organisation or person associated with?
    options:
      - Agriculture, Forestry and Fishing
      - Mining and Quarrying
      - Manufacturing
      - Electricity, gas, steam and air conditioning supply
      - Water supply, sewerage, waste management and remediation activities
      - Construction
      - Wholesale and retail trade; repair of motor vehicles and motorcycles
      - Transportation and storage
      - Accommodation and food service activities
      - Information and communication
      - Financial and insurance activities
      - Real estate activities
      - Professional, scientific and technical activities
      - Administrative and support service activities
      - Public administration and defence; compulsory social security
      - Education
      - Human health and social work activities
      - Arts, entertainment and recreation
      - Political group
      - NGO or think-tank
      - Trade union
      - Other
      - Don't know
  explanation:
    type: str
    description: A few paragraphs explaining your decision in English, formatted in Markdown. In the explanation, link to the most relevant sources from the provided documents. Include at least one inline URL.
  confidence:
    type: int
    description: Your confidence level in the decision, on a scale of 1 (lowest) to 10 (highest). If you don't have enough information or the documents refer to different organisations that may share a name, please set this to 1.
examples:
  - input: "Charles A Daniel-Hobbs"
    output:
      industry: Financial and insurance activities
      explanation: |
        According to [the Wall Street Journal](https://www.wsj.com/market-data/quotes/SFNC/company-people/executive-profile/247375783), Mr. Charles Alexander DANIEL-HOBBS is the Chief Financial Officer and Executive Vice President of Simmons First National Corp, a bank holding company.
        
        A Charles Alexander DANIEL-HOBBS also operates several companies, such as [DIBDEN PROPERTY LIMITED](https://find-and-update.company-information.service.gov.uk/company/10126637), which Companies House classifies as "Other letting and operating of own or leased real estate". However, the information is not clear on whether these are the same person.
      confidence: 2
  - input: "Unite the Union"
    output:
      industry: Trade union
      explanation: |
        Unite is [one of the two largest trade unions in the UK](https://en.wikipedia.org/wiki/Unite_the_Union), with over 1.2 million members. It represents various industries, such as construction, manufacturing, transport, logistics and other sectors.
      confidence: 7
  - input: "Google UK Limited"
    output:
      industry: Information and communication
      explanation: |
        Google UK Limited is a [subsidiary of Google LLC](https://about.google/intl/ALL_uk/google-in-uk/), a multinational technology company that specializes in Internet-related services and products.

        The company [provides various web based business services](https://www.bloomberg.com/profile/company/1200719Z:LN), including a web based search engine which includes various options such as web, image, directory, and news searches. 
      confidence: 10
  - input: "John Smith"
    output:
      industry: Don't know
      explanation: |
        The documents about John Smith refer to multiple people, so there's no way to accurately assess what industry this particular individual belongs to.
      confidence: 1
```

A few things to note about the configuration:

- We're using Brave due to its generous free API tier, but Google search results are generally more accurate.
- Because our dataset is UK-centric, we're setting the country to "GB" in Brave. This should help filter out some of the irrelevant results.
- The industries are based on the [Standard Industrial Classification](https://resources.companieshouse.gov.uk/sic/) groups. You can probably come up with something more clever.
- We are using [chain-of-thought](https://github.com/Global-Witness/augmenta/blob/main/docs/prompt.md#use-chain-of-thoguht) to guide the model through the process of classification. This can help improve accuracy.

Because we're using Brave and an OpenAI model, we need API keys for both services. Save them to a file called `.env` in the root of your project directory:

```
BRAVE_API_KEY=YOUR_KEY_GOES_HERE
OPENAI_API_KEY=YOUR_KEY_GOES_HERE
```

## Running the augmentation

Make sure you have `augmenta` [installed](https://github.com/Global-Witness/augmenta/tree/main?tab=readme-ov-file#installation), open the terminal and navigate to the directory where you saved the data, configuration file and API keys.

Run the following command to start the classification.

```bash
augmenta config.yaml
```

This should take a few seconds. Once it's done, you'll have a new file called `data/donations_classified.csv` with the augmented data.

```{r}
library(gt)

read_csv("data/donations_classified.csv") |>
  select(-DonorStatus, -CompanyRegistrationNumber, -chain_of_thought) |>
  gt() |>
  fmt_markdown(columns = "explanation")
```

## Results

There are a few things to note here.

First, the results are only as good [as the information they're fed](https://en.wikipedia.org/wiki/Garbage_in%2C_garbage_out). Google search results tend to be better than those offered by Brave or Duckduckgo, but they're not perfect either.

This is particularly an issue with individuals with generic names. We have instructed the model to flag those cases as "Don't know", but it can still lead to some issues.

For example, the "Robert H Miall" identified in the search results has passed away in 2011 (as noted by the LLM itself), so he can't be the donor we're looking for (the donations are from 2024).

We can work around these limitations in a few ways:

- Increase the number of results to get a better picture of the donor.
- Use a better search engine and/or more specific search query.
- Be more descriptive about these edge cases in the prompt and examples.
- Filter out individual donors and stick to organisations.

For organisations, which tend to be easier to identify, the model does a much better job.

If we were to publish any analysis of this data, we would need to fact-check the results.